---
editor_options: 
  markdown: 
    wrap: 100
---

## Useful links:

Conversions: Seurat -\> h5Seurat -\> AnnData
<https://mojaveazure.github.io/seurat-disk/articles/convert-anndata.html>

• Ontologies: <https://www.ebi.ac.uk/ols/>

schema_guide.md · chanzuckerberg/single-cell-curation
<https://github.com/chanzuckerberg/single-cell-curation/blob/main/docs/schema_guide.md>

corpora_schema.md at chanzuckerberg/single-cell-curation
<https://github.com/chanzuckerberg/single-cell-curation/blob/main/docs/corpora_schema.md>

[anndata 0.7.6 documentation](https://anndata.readthedocs.io/en/latest/)

## Install required packages:

Install hdf5 python package.

```{bash}
yay hdf5 # installed first option
```

Install SeuratDisk.

```{r}
remotes::install_github("mojaveazure/seurat-disk")
remotes::update_packages('Seurat')
```

## Set up stuff:

```{r}
setwd("~/cortical_arealization/browser")

p_load(SeuratDisk)
p_load(tidyverse)
```

## Make smaller subset of neocortex object:

```{r}

load("../neocortex/data/cfc4b2_neocortex_v3.RData") # 94GB


neocortex_subset <- subsetSeurat(Neocortex, 
                                 fraction = 0.01, 
                                 write.to = "neocortex/data/210318_neocortex_v3_subset.01.rds")

write_rds(neocortex_subset, "neocortex/data/210318_neocortex_v3_subset.01.rds")                   
write_rds(Neocortex, "neocortex/data/210318_neocortex_v3.rds")

# Update Neocortex to Seurat v4:
neocortex <- UpdateSeuratObject(Neocortex)
rm(Neocortex)
write_rds(neocortex, '../neocortex/data/neocortex_v4.rds')
```

```{r}
neocortex <- read_rds("../neocortex/data/neocortex_v3_updated.rds")
SaveH5Seurat(object = neocortex, filename = "210318_neocortex")

```

```{r}
LoadH5Seurat("neocortex.h5seurat") 
# Error in `[[.H5Group`(assay.group, "features") : An object with name features does not exist in this group


hfile <- Connect("neocortex.h5Seurat")
hfile$index()
# Data for assay RNA★ (default assay)
#   counts      data    scale.data
#     ✔          ✔          ✖     
```

```{r}
neocortex_subset <- subsetSeurat(neocortex, fraction = 0.01, write.to = "../neocortex/data/210318_neocortex_subset0.01.rds")
neocortex_subset <- read_rds("../neocortex/data/210318_neocortex_subset0.01.rds")
```

```{r}
An object of class Seurat 
29805 features across 4042 samples within 1 assay 
Active assay: RNA (29805 features, 2000 variable features)
 3 dimensional reductions calculated: fastPCA, pca, umap

Formal class 'Seurat' [package "SeuratObject"] with 13 slots
  ..@ assays      :List of 1
  .. ..$ RNA:Formal class 'Assay' [package "SeuratObject"] with 8 slots
  ..@ meta.data   :'data.frame':        4042 obs. of  15 variables:
  .. ..$ nGene            : int [1:4042] 1144 2853 1157 4288 1189 831 2007 1743 1265 2352 ...
  .. ..$ nUMI             : num [1:4042] 2015 7260 2533 14590 2545 ...
  .. ..$ orig.ident       : Factor w/ 1 level "SeuratProject": 1 1 1 1 1 1 1 1 1 1 ...
  .. ..$ Age              : num [1:4042] 20 19 18 25 20 25 25 19 18 19 ...
  .. ..$ Individual       : chr [1:4042] "GW20_34" "GW19" "GW18" "GW25" ...
  .. ..$ Area             : chr [1:4042] "PFC" "somatosensory" "motor" "parietal" ...
  .. ..$ Area.sub         : chr [1:4042] "NA" "NA" "NA" "NA" ...
  .. ..$ Structure        : chr [1:4042] "neocortex" "Neocortex" "neocortex" "neocortex" ...
  .. ..$ Lamina           : chr [1:4042] "VZ" "all" "all" "VZ" ...
  .. ..$ percent.mito     : num [1:4042] 4.26e-16 3.72e-44 1.19e-27 3.86e-97 2.70e-46 ...
  .. ..$ percent.ribo     : num [1:4042] 0.207 0.221 0.204 0.159 0.312 ...
  .. ..$ V1               : int [1:4042] 36 28 28 23 10 26 13 15 5 6 ...
  .. ..$ nCount_RNA       : num [1:4042] 61290 190614 59044 317943 81847 ...
  .. ..$ nFeature_RNA     : int [1:4042] 1144 2853 1157 4288 1189 831 2007 1743 1263 2352 ...
  .. ..$ ConsensusCellType: Factor w/ 12 levels "0","CR","Dividing",..: 4 3 8 6 4 6 6 11 1 3 ...
  ..@ active.assay: chr "RNA"
  ..@ active.ident: Factor w/ 47 levels "3","4","5","6",..: 34 26 26 21 8 24 11 13 3 4 ...
  .. ..- attr(*, "names")= chr [1:4042] "GW2034_PFCVZ_TTGTAGGGTTACAGAA" "gw19_2_SS_TGGGATTGTCCATCTC" "GW18_motor_TTAGGCAGTGTAACGG" "GW25_parietalVZ_CTAGTGAAGCGCTCCA" ...
  ..@ graphs      : list()
  ..@ neighbors   : list()
  ..@ reductions  :List of 3
  .. ..$ fastPCA:Formal class 'DimReduc' [package "SeuratObject"] with 9 slots
  .. ..$ pca    :Formal class 'DimReduc' [package "SeuratObject"] with 9 slots
  .. ..$ umap   :Formal class 'DimReduc' [package "SeuratObject"] with 9 slots
  ..@ images      : list()
  ..@ project.name: chr "SeuratProject"
  ..@ misc        : list()
  ..@ version     :Classes 'package_version', 'numeric_version'  hidden list of 1
  .. ..$ : int [1:3] 4 0 1
  ..@ commands    :List of 4
  .. ..$ FindVariableFeatures.RNA:Formal class 'SeuratCommand' [package "SeuratObject"] with 5 slots
  .. ..$ ScaleData.RNA           :Formal class 'SeuratCommand' [package "SeuratObject"] with 5 slots
  .. ..$ RunPCA.RNA              :Formal class 'SeuratCommand' [package "SeuratObject"] with 5 slots
  .. ..$ RunUMAP.RNA.pca         :Formal class 'SeuratCommand' [package "SeuratObject"] with 5 slots
  ..@ tools       : list()
```

```{r}
neocortex_subset@assays$RNA@key <- "rna_"
neocortex_subset@reductions$fastPCA@key <- 'fastPC_'   
neocortex_subset@reductions <- neocortex_subset@reductions[2:3]

tic("SaveH5Seurat")
SaveH5Seurat(object = neocortex_subset, filename = "210318_neocortex_subset", overwrite = TRUE)
toc()
hfile <- Connect("210318_neocortex_subset.h5seurat")
hfile$index()
LoadH5Seurat("210318_neocortex_subset.h5seurat")
```

```{r}
p_load(SeuratData)
data("pbmc3k.final")
pbmc3k.final@assays$RNA %>% str
```

## Formal class 'Seurat' [package "Seurat"] with 13 slots

..\@ assays :List of 1 .. ..\$ RNA:Formal class 'Assay' [package "Seurat"] with 8 slots .. .. ..
..\@ counts :Formal class 'dgCMatrix' [package "Matrix"] with 6 slots .. .. .. .. .. ..\@ i : int
[1:2238732] 29 73 80 148 163 184 186 227 229 230 ... .. .. .. .. .. ..\@ p : int [1:2639] 0 779 2131
3260 4220 4741 5522 6304 7094 7626 ... .. .. .. .. .. ..\@ Dim : int [1:2] 13714 2638 .. .. .. .. ..
..\@ Dimnames:List of 2 .. .. .. .. .. .. ..\$ : chr [1:13714] "AL627309.1" "AP006222.2"
"RP11-206L10.2" "RP11-206L10.9" ... .. .. .. .. .. .. ..\$ : chr [1:2638] "AAACATACAACCAC"
"AAACATTGAGCTAC" "AAACATTGATCAGC" "AAACCGTGCTTCCG" ... .. .. .. .. .. ..\@ x : num [1:2238732] 1 1 2
1 1 1 1 41 1 1 ... .. .. .. .. .. ..\@ factors : list() .. .. .. ..\@ data :Formal class 'dgCMatrix'
[package "Matrix"] with 6 slots .. .. .. .. .. ..\@ i : int [1:2238732] 29 73 80 148 163 184 186 227
229 230 ... .. .. .. .. .. ..\@ p : int [1:2639] 0 779 2131 3260 4220 4741 5522 6304 7094 7626 ...
.. .. .. .. .. ..\@ Dim : int [1:2] 13714 2638 .. .. .. .. .. ..\@ Dimnames:List of 2 .. .. .. .. ..
.. ..\$ : chr [1:13714] "AL627309.1" "AP006222.2" "RP11-206L10.2" "RP11-206L10.9" ... .. .. .. .. ..
.. ..\$ : chr [1:2638] "AAACATACAACCAC" "AAACATTGAGCTAC" "AAACATTGATCAGC" "AAACCGTGCTTCCG" ... .. ..
.. .. .. ..\@ x : num [1:2238732] 1.64 1.64 2.23 1.64 1.64 ... .. .. .. .. .. ..\@ factors : list()
.. .. .. ..\@ scale.data : num [1:13714, 1:2638] -0.0581 -0.0336 -0.0417 -0.0336 -0.0822 ... .. ..
.. .. ..- attr(*, "dimnames")=List of 2 .. .. .. .. .. ..\$ : chr [1:13714] "AL627309.1"
"AP006222.2" "RP11-206L10.2" "RP11-206L10.9" ... .. .. .. .. .. ..\$ : chr [1:2638] "AAACATACAACCAC"
"AAACATTGAGCTAC" "AAACATTGATCAGC" "AAACCGTGCTTCCG" ... .. .. .. ..\@ key : chr "rna\_" .. .. .. ..\@
assay.orig : NULL .. .. .. ..\@ var.features : chr [1:2000] "PPBP" "LYZ" "S100A9" "IGLL5" ... .. ..
.. ..\@ meta.features:'data.frame': 13714 obs. of 5 variables: .. .. .. .. ..\$ vst.mean : num
[1:13714] 0.00341 0.00114 0.0019 0.00114 0.00682 ... .. .. .. .. ..\$ vst.variance : num [1:13714]
0.0034 0.00114 0.00189 0.00114 0.00678 ... .. .. .. .. ..\$ vst.variance.expected : num [1:13714]
0.00365 0.00114 0.00197 0.00114 0.00748 ... .. .. .. .. ..\$ vst.variance.standardized: num
[1:13714] 0.933 0.992 0.963 0.992 0.906 ... .. .. .. .. ..\$ vst.variable : logi [1:13714] FALSE
FALSE FALSE FALSE FALSE FALSE ... .. .. .. ..\@ misc : NULL ..\@ meta.data :'data.frame': 2638 obs.
of 7 variables: .. ..\$ orig.ident : Factor w/ 1 level "pbmc3k": 1 1 1 1 1 1 1 1 1 1 ... .. ..\$
nCount_RNA : num [1:2638] 2419 4903 3147 2639 980 ... .. ..\$ nFeature_RNA : int [1:2638] 779 1352
1129 960 521 781 782 790 532 550 ... .. ..\$ seurat_annotations: Factor w/ 9 levels "Naive CD4
T",..: 2 4 2 3 7 2 5 5 1 6 ... .. ..\$ percent.mt : num [1:2638] 3.02 3.79 0.89 1.74 1.22 ... ..
..\$ RNA_snn_res.0.5 : Factor w/ 9 levels "0","1","2","3",..: 2 4 2 3 7 2 5 5 1 6 ... .. ..\$
seurat_clusters : Factor w/ 9 levels "0","1","2","3",..: 2 4 2 3 7 2 5 5 1 6 ... ..\@ active.assay:
chr "RNA" ..\@ active.ident: Factor w/ 9 levels "Naive CD4 T",..: 2 4 2 3 7 2 5 5 1 6 ... .. ..-
attr(*, "names")= chr [1:2638] "AAACATACAACCAC" "AAACATTGAGCTAC" "AAACATTGATCAGC" "AAACCGTGCTTCCG"
... ..\@ graphs :List of 2 .. ..\$ RNA_nn :Formal class 'Graph' [package "Seurat"] with 7 slots ..
.. .. ..\@ assay.used: chr "RNA" .. .. .. ..\@ i : int [1:52760] 0 6 102 203 292 421 451 511 547 618
... .. .. .. ..\@ p : int [1:2639] 0 37 55 78 95 102 113 150 173 184 ... .. .. .. ..\@ Dim : int
[1:2] 2638 2638 .. .. .. ..\@ Dimnames :List of 2 .. .. .. .. ..\$ : chr [1:2638] "AAACATACAACCAC"
"AAACATTGAGCTAC" "AAACATTGATCAGC" "AAACCGTGCTTCCG" ... .. .. .. .. ..\$ : chr [1:2638]
"AAACATACAACCAC" "AAACATTGAGCTAC" "AAACATTGATCAGC" "AAACCGTGCTTCCG" ... .. .. .. ..\@ x : num
[1:52760] 1 1 1 1 1 1 1 1 1 1 ... .. .. .. ..\@ factors : list() .. ..\$ RNA_snn:Formal class
'Graph' [package "Seurat"] with 7 slots .. .. .. ..\@ assay.used: chr "RNA" .. .. .. ..\@ i : int
[1:194704] 0 6 76 102 167 171 187 203 229 292 ... .. .. .. ..\@ p : int [1:2639] 0 80 145 197 252
299 367 463 523 573 ... .. .. .. ..\@ Dim : int [1:2] 2638 2638 .. .. .. ..\@ Dimnames :List of 2 ..
.. .. .. ..\$ : chr [1:2638] "AAACATACAACCAC" "AAACATTGAGCTAC" "AAACATTGATCAGC" "AAACCGTGCTTCCG" ...
.. .. .. .. ..\$ : chr [1:2638] "AAACATACAACCAC" "AAACATTGAGCTAC" "AAACATTGATCAGC" "AAACCGTGCTTCCG"
... .. .. .. ..\@ x : num [1:194704] 1 0.1111 0.0811 0.2121 0.0811 ... .. .. .. ..\@ factors :
list() ..\@ neighbors : list() ..\@ reductions :List of 2 .. ..\$ pca :Formal class 'DimReduc'
[package "Seurat"] with 9 slots .. .. .. ..\@ cell.embeddings : num [1:2638, 1:50] -4.73 -0.517
-3.189 12.793 -3.129 ... .. .. .. .. ..- attr(*, "dimnames")=List of 2 .. .. .. .. .. ..\$ : chr
[1:2638] "AAACATACAACCAC" "AAACATTGAGCTAC" "AAACATTGATCAGC" "AAACCGTGCTTCCG" ... .. .. .. .. .. ..\$
: chr [1:50] "PC_1" "PC_2" "PC_3" "PC_4" ... .. .. .. ..\@ feature.loadings : num [1:2000, 1:50]
0.01099 0.11623 0.11541 -0.00799 -0.01524 ... .. .. .. .. ..- attr(*, "dimnames")=List of 2 .. .. ..
.. .. ..\$ : chr [1:2000] "PPBP" "LYZ" "S100A9" "IGLL5" ... .. .. .. .. .. ..\$ : chr [1:50] "PC_1"
"PC_2" "PC_3" "PC_4" ... .. .. .. ..\@ feature.loadings.projected: num[0 , 0 ] .. .. .. ..\@
assay.used : chr "RNA" .. .. .. ..\@ global : logi FALSE .. .. .. ..\@ stdev : num [1:50] 7.1 4.5
3.87 3.75 3.17 ... .. .. .. ..\@ key : chr "PC\_" .. .. .. ..\@ jackstraw :Formal class
'JackStrawData' [package "Seurat"] with 4 slots .. .. .. .. .. ..\@ empirical.p.values : num
[1:2000, 1:20] 0 0 0 0.0075 0 0 0.008 0 0.016 0 ... .. .. .. .. .. .. ..- attr(*, "dimnames")=List
of 2 .. .. .. .. .. .. .. ..\$ : chr [1:2000] "PPBP" "LYZ" "S100A9" "IGLL5" ... .. .. .. .. .. .. ..
..\$ : chr [1:20] "PC1" "PC2" "PC3" "PC4" ... .. .. .. .. .. ..\@ fake.reduction.scores : num
[1:2000, 1:20] -0.003133 0.000127 -0.000119 0.002002 0.000613 ... .. .. .. .. .. ..\@
empirical.p.values.full: logi [1, 1] NA .. .. .. .. .. ..\@ overall.p.values : num [1:20, 1:2] 1 2 3
4 5 6 7 8 9 10 ... .. .. .. .. .. .. ..- attr(*, "dimnames")=List of 2 .. .. .. .. .. .. .. ..\$ :
NULL .. .. .. .. .. .. .. ..\$ : chr [1:2] "PC" "Score" .. .. .. ..\@ misc :List of 1 .. .. .. ..
..\$ total.variance: num 1734 .. ..\$ umap:Formal class 'DimReduc' [package "Seurat"] with 9 slots
.. .. .. ..\@ cell.embeddings : num [1:2638, 1:2] -4.23 -4.89 -5.51 11.33 -7.45 ... .. .. .. .. ..-
attr(*, "scaled:center")= num [1:2] 1.964 -0.201 .. .. .. .. ..- attr(*, "dimnames")=List of 2 .. ..
.. .. .. ..\$ : chr [1:2638] "AAACATACAACCAC" "AAACATTGAGCTAC" "AAACATTGATCAGC" "AAACCGTGCTTCCG" ...
.. .. .. .. .. ..\$ : chr [1:2] "UMAP_1" "UMAP_2" .. .. .. ..\@ feature.loadings : num[0 , 0 ] .. ..
.. ..\@ feature.loadings.projected: num[0 , 0 ] .. .. .. ..\@ assay.used : chr "RNA" .. .. .. ..\@
global : logi TRUE .. .. .. ..\@ stdev : num(0) .. .. .. ..\@ key : chr "UMAP\_" .. .. .. ..\@
jackstraw :Formal class 'JackStrawData' [package "Seurat"] with 4 slots .. .. .. .. .. ..\@
empirical.p.values : num[0 , 0 ] .. .. .. .. .. ..\@ fake.reduction.scores : num[0 , 0 ] .. .. .. ..
.. ..\@ empirical.p.values.full: num[0 , 0 ] .. .. .. .. .. ..\@ overall.p.values : num[0 , 0 ] ..
.. .. ..\@ misc : list() ..\@ project.name: chr "pbmc3k" ..\@ misc : list() ..\@ version :Classes
'package_version', 'numeric_version' hidden list of 1 .. ..\$ : int [1:3] 3 1 4 ..\@ commands :List
of 9 .. ..\$ NormalizeData.RNA :Formal class 'SeuratCommand' [package "Seurat"] with 5 slots .. ..
.. ..\@ name : chr "NormalizeData.RNA" .. .. .. ..\@ time.stamp : POSIXct[1:1], format: "2020-04-29
21:50:43" .. .. .. ..\@ assay.used : chr "RNA" .. .. .. ..\@ call.string: chr
"NormalizeData(pbmc3k.final)" .. .. .. ..\@ params :List of 5 .. .. .. .. ..\$ assay : chr "RNA" ..
.. .. .. ..\$ normalization.method: chr "LogNormalize" .. .. .. .. ..\$ scale.factor : num 10000 ..
.. .. .. ..\$ margin : num 1 .. .. .. .. ..\$ verbose : logi TRUE .. ..\$
FindVariableFeatures.RNA:Formal class 'SeuratCommand' [package "Seurat"] with 5 slots .. .. .. ..\@
name : chr "FindVariableFeatures.RNA" .. .. .. ..\@ time.stamp : POSIXct[1:1], format: "2020-04-29
21:50:45" .. .. .. ..\@ assay.used : chr "RNA" .. .. .. ..\@ call.string: chr
"FindVariableFeatures(pbmc3k.final)" .. .. .. ..\@ params :List of 12 .. .. .. .. ..\$ assay : chr
"RNA" .. .. .. .. ..\$ selection.method : chr "vst" .. .. .. .. ..\$ loess.span : num 0.3 .. .. ..
.. ..\$ clip.max : chr "auto" .. .. .. .. ..\$ mean.function :function (mat, display_progress)\
.. .. .. .. ..\$ dispersion.function:function (mat, display_progress)\
.. .. .. .. ..\$ num.bin : num 20 .. .. .. .. ..\$ binning.method : chr "equal_width" .. .. .. ..
..\$ nfeatures : num 2000 .. .. .. .. ..\$ mean.cutoff : num [1:2] 0.1 8 .. .. .. .. ..\$
dispersion.cutoff : num [1:2] 1 Inf .. .. .. .. ..\$ verbose : logi TRUE .. ..\$ ScaleData.RNA
:Formal class 'SeuratCommand' [package "Seurat"] with 5 slots .. .. .. ..\@ name : chr
"ScaleData.RNA" .. .. .. ..\@ time.stamp : POSIXct[1:1], format: "2020-04-29 21:50:47" .. .. .. ..\@
assay.used : chr "RNA" .. .. .. ..\@ call.string: chr "ScaleData(pbmc3k.final, features =
rownames(pbmc3k.final))" .. .. .. ..\@ params :List of 10 .. .. .. .. ..\$ features : chr [1:13714]
"AL627309.1" "AP006222.2" "RP11-206L10.2" "RP11-206L10.9" ... .. .. .. .. ..\$ assay : chr "RNA" ..
.. .. .. ..\$ model.use : chr "linear" .. .. .. .. ..\$ use.umi : logi FALSE .. .. .. .. ..\$
do.scale : logi TRUE .. .. .. .. ..\$ do.center : logi TRUE .. .. .. .. ..\$ scale.max : num 10 ..
.. .. .. ..\$ block.size : num 1000 .. .. .. .. ..\$ min.cells.to.block: num 2638 .. .. .. .. ..\$
verbose : logi TRUE .. ..\$ RunPCA.RNA :Formal class 'SeuratCommand' [package "Seurat"] with 5 slots
.. .. .. ..\@ name : chr "RunPCA.RNA" .. .. .. ..\@ time.stamp : POSIXct[1:1], format: "2020-04-29
21:50:54" .. .. .. ..\@ assay.used : chr "RNA" .. .. .. ..\@ call.string: chr "RunPCA(pbmc3k.final,
features = VariableFeatures(pbmc3k.final))" .. .. .. ..\@ params :List of 11 .. .. .. .. ..\$ assay
: chr "RNA" .. .. .. .. ..\$ features : chr [1:2000] "PPBP" "LYZ" "S100A9" "IGLL5" ... .. .. .. ..
..\$ npcs : num 50 .. .. .. .. ..\$ rev.pca : logi FALSE .. .. .. .. ..\$ weight.by.var : logi TRUE
.. .. .. .. ..\$ verbose : logi TRUE .. .. .. .. ..\$ ndims.print : int [1:5] 1 2 3 4 5 .. .. .. ..
..\$ nfeatures.print: num 30 .. .. .. .. ..\$ reduction.name : chr "pca" .. .. .. .. ..\$
reduction.key : chr "PC\_" .. .. .. .. ..\$ seed.use : num 42 .. ..\$ JackStraw.RNA.pca :Formal
class 'SeuratCommand' [package "Seurat"] with 5 slots .. .. .. ..\@ name : chr "JackStraw.RNA.pca"
.. .. .. ..\@ time.stamp : POSIXct[1:1], format: "2020-04-29 21:54:47" .. .. .. ..\@ assay.used :
chr "RNA" .. .. .. ..\@ call.string: chr "JackStraw(pbmc3k.final)" .. .. .. ..\@ params :List of 7
.. .. .. .. ..\$ reduction : chr "pca" .. .. .. .. ..\$ assay : chr "RNA" .. .. .. .. ..\$ dims :
num 20 .. .. .. .. ..\$ num.replicate: num 100 .. .. .. .. ..\$ prop.freq : num 0.01 .. .. .. ..
..\$ verbose : logi TRUE .. .. .. .. ..\$ maxit : num 1000 .. ..\$ ScoreJackStraw :Formal class
'SeuratCommand' [package "Seurat"] with 5 slots .. .. .. ..\@ name : chr "ScoreJackStraw" .. .. ..
..\@ time.stamp : POSIXct[1:1], format: "2020-04-29 21:54:50" .. .. .. ..\@ assay.used : chr "RNA"
.. .. .. ..\@ call.string: chr "ScoreJackStraw(pbmc3k.final, dims = 1:20)" .. .. .. ..\@ params
:List of 4 .. .. .. .. ..\$ reduction : chr "pca" .. .. .. .. ..\$ dims : int [1:20] 1 2 3 4 5 6 7 8
9 10 ... .. .. .. .. ..\$ score.thresh: num 1e-05 .. .. .. .. ..\$ do.plot : logi FALSE .. ..\$
FindNeighbors.RNA.pca :Formal class 'SeuratCommand' [package "Seurat"] with 5 slots .. .. .. ..\@
name : chr "FindNeighbors.RNA.pca" .. .. .. ..\@ time.stamp : POSIXct[1:1], format: "2020-04-29
21:54:51" .. .. .. ..\@ assay.used : chr "RNA" .. .. .. ..\@ call.string: chr
"FindNeighbors(pbmc3k.final, dims = 1:10)" .. .. .. ..\@ params :List of 13 .. .. .. .. ..\$
reduction : chr "pca" .. .. .. .. ..\$ dims : int [1:10] 1 2 3 4 5 6 7 8 9 10 .. .. .. .. ..\$ assay
: chr "RNA" .. .. .. .. ..\$ k.param : num 20 .. .. .. .. ..\$ compute.SNN : logi TRUE .. .. .. ..
..\$ prune.SNN : num 0.0667 .. .. .. .. ..\$ nn.method : chr "rann" .. .. .. .. ..\$ annoy.metric:
chr "euclidean" .. .. .. .. ..\$ nn.eps : num 0 .. .. .. .. ..\$ verbose : logi TRUE .. .. .. ..
..\$ force.recalc: logi FALSE .. .. .. .. ..\$ do.plot : logi FALSE .. .. .. .. ..\$ graph.name :
chr [1:2] "RNA_nn" "RNA_snn" .. ..\$ FindClusters :Formal class 'SeuratCommand' [package "Seurat"]
with 5 slots .. .. .. ..\@ name : chr "FindClusters" .. .. .. ..\@ time.stamp : POSIXct[1:1],
format: "2020-04-29 21:54:53" .. .. .. ..\@ assay.used : chr "RNA" .. .. .. ..\@ call.string: chr
"FindClusters(pbmc3k.final, resolution = 0.5)" .. .. .. ..\@ params :List of 10 .. .. .. .. ..\$
graph.name : chr "RNA_snn" .. .. .. .. ..\$ modularity.fxn : num 1 .. .. .. .. ..\$ resolution : num
0.5 .. .. .. .. ..\$ method : chr "matrix" .. .. .. .. ..\$ algorithm : num 1 .. .. .. .. ..\$
n.start : num 10 .. .. .. .. ..\$ n.iter : num 10 .. .. .. .. ..\$ random.seed : num 0 .. .. .. ..
..\$ group.singletons: logi TRUE .. .. .. .. ..\$ verbose : logi TRUE .. ..\$ RunUMAP.RNA.pca
:Formal class 'SeuratCommand' [package "Seurat"] with 5 slots .. .. .. ..\@ name : chr
"RunUMAP.RNA.pca" .. .. .. ..\@ time.stamp : POSIXct[1:1], format: "2020-04-29 21:55:01" .. .. ..
..\@ assay.used : chr "RNA" .. .. .. ..\@ call.string: chr "RunUMAP(pbmc3k.final, dims = 1:10)" ..
.. .. ..\@ params :List of 20 .. .. .. .. ..\$ dims : int [1:10] 1 2 3 4 5 6 7 8 9 10 .. .. .. ..
..\$ reduction : chr "pca" .. .. .. .. ..\$ assay : chr "RNA" .. .. .. .. ..\$ umap.method : chr
"uwot" .. .. .. .. ..\$ n.neighbors : int 30 .. .. .. .. ..\$ n.components : int 2 .. .. .. .. ..\$
metric : chr "cosine" .. .. .. .. ..\$ learning.rate : num 1 .. .. .. .. ..\$ min.dist : num 0.3 ..
.. .. .. ..\$ spread : num 1 .. .. .. .. ..\$ set.op.mix.ratio : num 1 .. .. .. .. ..\$
local.connectivity : int 1 .. .. .. .. ..\$ repulsion.strength : num 1 .. .. .. .. ..\$
negative.sample.rate: int 5 .. .. .. .. ..\$ uwot.sgd : logi FALSE .. .. .. .. ..\$ seed.use : int
42 .. .. .. .. ..\$ angular.rp.forest : logi FALSE .. .. .. .. ..\$ verbose : logi TRUE .. .. .. ..
..\$ reduction.name : chr "umap" .. .. .. .. ..\$ reduction.key : chr "UMAP\_" ..\@ tools : list()
..\@ NA : NULL

## Update pbmc3k

```{r}
pbmc3k.final_update <- UpdateSeuratObject(pbmc3k.final)
```

    Validating object structure
    Updating object slots
    Ensuring keys are in the proper strucutre
    Ensuring feature names don't have underscores or pipes
    Object representation is consistent with the most current Seurat version
    Not a validObject(): no slot of name "images" for this object of class "Seurat" 

```{r}
tic("SaveH5Seurat")
SaveH5Seurat(object = pbmc3k.final_update, filename = "210318_neocortex_subset", overwrite = TRUE)
toc()
hfile <- Connect("210318_neocortex_subset.h5seurat")
hfile$index()
```

# gw16_neocortex

```{r}
gw16_neocortex <- read_rds(path="neocortex/data/gw16_neocortex_v3.rds")

SaveH5Seurat(object = gw16_neocortex, filename = "gw16_neocortex")
```

    Creating h5Seurat file for version 3.1.5.9900
    Error in attr$write(robj) : HDF5-API Errors:
        error #000: H5A.c in H5Awrite(): line 676: buf parameter can't be NULL
            class: HDF5
            major: Invalid arguments to routine
            minor: Bad value

    LoadH5Seurat("gw16_neocortex.h5seurat") # Cannot find dataset with cell names

```{r}
load("../neocortex/data/gw16neo.RData")
 gw16neo@version
# [1] ‘2.3.0’
SaveH5Seurat(object = gw16neo, filename = "gw16_neo")
# Error: Unable to coerce an object of class seurat to a Seurat object
gw16_neocortex <- UpdateSeuratObject(gw16neo)

# Updating from v2.X to v3.X
# Error in `[[<-`(`*tmp*`, names(x = n.calc), value = n.calc) : 
#   [[<- defined for objects of type "S4" only for subclasses of environment

# https://github.com/satijalab/seurat/issues/3964#issuecomment-802417912
Updating from v2.X to v3.X
Validating object structure
Updating object slots
Ensuring keys are in the proper strucutre
Ensuring feature names don't have underscores or pipes
Object representation is consistent with the most current Seurat version

SaveH5Seurat(object = gw16_neocortex, filename = "210318_gw16_neocortex", overwrite = TRUE)
```

```{r}
hfile <- Connect("210318_gw16_neocortex.h5seurat")
hfile$index()
```

```{r}
Convert("210318_gw16_neocortex.h5seurat", dest = "h5ad")
```

```{r}
a <- LoadH5Seurat("210318_gw16_neocortex.h5seurat")
```

```{python}
import scanpy
adata = scanpy.read_h5ad("./browser-dbx/data/210323_neocortex_v3.2.3_subset.05.h5ad")
```

File "/home/csandoval/.local/lib/python3.9/site-packages/anndata/\_io/h5ad.py", line 442, in
read_h5ad return AnnData(\*\*d) TypeError: **init**() got an unexpected keyword argument
'active.ident'

```{r}
neocortex <- read_rds("../neocortex/data/210318_neocortex_v3.rds")
sms("Loaded neocortex")

neocortex@version
# 3.1.0
# Error: Object too old to save, please update your Seurat object to at least v3.1.2 using UpdateSeuratObject

neocortex <- UpdateSeuratObject(neocortex)
write_rds(neocortex, "../neocortex/data/210318_neocortex_v3.2.3.rds")
sms(paste(Sys.time(), "write_rds neocortex v3.2.3"))


```

```{r}
gw16 <- read_rds("../neocortex/data/gw16_neocortex_v3.rds")
gw16@version # 4.0.1

load("../neocortex/data/gw16neo.RData")
# 2.3.0
write_rds(neocortex, "../neocortex/data/210318_neocortex_v3.2.3.rds")

gw16 <- read_rds("../neocortex/data/210322_gw16_neocortex_v2.rds")
gw16 <- UpdateSeuratObject(gw16)
write_rds(gw16, "../neocortex/data/210322_gw16_neocortex_v3.2.3.rds")
```

## Do for gw19

```{r}
load("../neocortex/data/gw19_2.RData")
gw19_2@version #2.3.0
gw19_2 <- UpdateSeuratObject(gw19_2)

write_rds(gw19_2, "../neocortex/data/210322_gw19_2_v3.2.3.rds")
```

```{r}
gw16 <- read_rds("../neocortex/data/210322_gw16_neocortex_v3.2.3.rds")
gw19_2 <- read_rds("../neocortex/data/210322_gw19_2_v3.2.3.rds")

tic()
neocortex <- read_rds("../neocortex/data/210318_neocortex_v3.2.3.rds")
push.toc("Neocortex loaded")
gw16 <- read_rds("../neocortex/data/210322_gw16_neocortex_v3.2.3.rds")
gw19_2 <- read_rds("../neocortex/data/210322_gw19_2_v3.2.3.rds")
push.toc("gw16 and gw19 loaded")
```

```{r}
tic()
neocortex_full <- merge(x = neocortex,  y = gw16)
rm(neocortex)
write_rds(neocortex_full, "../neocortex/data/210322_neocortex_full_v3.2.3.rds")
push.toc("Mege all")
```

```{r}
neocortex_path <- "../neocortex/data/210322_neocortex_full_v3.2.3.rds"
neocortex <- read_rds(neocortex_path) # has gw16 cells
# 457, 965 cells
```

```{r}
# TODO NEEDS TO INCLUDE SCALED DATA WHEN MERGING GW16.
# Change individual to gw19_2

neocortex_subset <- subsetSeurat(neocortex, fraction = 0.01, 
                                 write.to = "../neocortex/data/210322_neocortex_full_v3.2.3_subset_0.01.rds")

tic()
write_rds(neocortex, path = neocortex_path)
toc()
sms("write_rds")

neocortex_subset@meta.data %>% head

# Run seurat_to_h5ad in a separate session (SeuratDisk causes problems with Seurat functions.)

meta.data %>% count(Individual, cluster_label, cell_type)
# Message to Aparna @ 2021-03-23
# 0 = outlier
# outlier only used in gw16
# other is from all samples except gw16

# I’m making sure that the types are all unified across samples - I’m including gw16 in the browser.
# Iwanted to ask if you think we should consolidate ‘Other’ with the ‘Outlier’ population, or leave 
# them as separate. As well as the ‘Mixed’ annotation.

# ‘mixed’ and ‘outlier’ were only used in gw16.
# Cells from all the other samples that were prev annotated as outliers were ‘0’ in the combo_2 column,
# so I binned those into the ‘outlier’ group.
# ‘Other’ was used in all samples except gw16.

```

# Restore scale.data and @reductions to neocortex object with all samples minus gw16 (unmerged)

```{r}
# No gw16 incorporated, but is v3.2.3
neocortex_old <- read_rds("../neocortex/data/210318_neocortex_v3.2.3.rds")

reductions <- neocortex_old@reductions
scale_data <- neocortex_old@assays$RNA@scale.data

# Remove gw16 cells from neocortex (full)
non_gw16_cells <- neocortex@meta.data %>% filter(!cell_id %>% str_detect("GW16")) %>% .$cell_id #53,747

neocortex_old <- subset(neocortex, cells = non_gw16_cells)
neocortex_old@reductions <- reductions
neocortex_old@assays$RNA@scale.data <- scale_data
```

## 2021-03-27

```{r}
# Merge neocortex v3 with gw16
tic()
neocortex_old <- read_rds("../neocortex/data/210318_neocortex_v3.2.3.rds") # No gw16 incorporated, but is v3.2.3
push.toc("Loaded")
gw16 <- read_rds("../neocortex/data/210322_gw16_neocortex_v3.2.3.rds")


tic()
neocortex_full <- merge(x = neocortex_old,  y = gw16, merge.data = TRUE)
toc("Merge") # 4m
tic()
write_rds(neocortex_full, "../neocortex/data/210322_neocortex_full_v3.2.3.rds")
toc("Save merged") # 8 minutes

# . . . . . . . . . . . . . . . . . . . . . . . . . . . . .
source("seurat_to_h5ad_func.R") # cleanMetadata, addReductionRows
neocortex_full_path <- "../neocortex/data/210322_neocortex_full_v3.2.3.rds"
neocortex_full <- read_rds(neocortex_full_path)

stash <- list(meta.data = neocortex_full@meta.data,
             reductions = neocortex_full@reductions)

# neocortex_full@meta.data <- stash$meta.data

# Clean metadata
neocortex_full@meta.data <- cleanMetadata(neocortex_full)
neocortex_full@meta.data %>% tail
# Add rows with reduction for gw16 NaN
gw16_cells <- neocortex_full@meta.data %>% filter(cell_id %>% str_detect("GW16")) %>% .$cell_id #53,747
reductions_tmp <- neocortex_old@reductions %>% map(~ addReductionRows(reduc_item = ., new_cells = gw16_cells) )
reductions_tmp %>% map(glimpse) # WHOHOOOOO!!! 457,000 rows
neocortex_full@reductions <- reductions_tmp
write_rds(neocortex_full, neocortex_full_path)
sms("Write_rds with clean metadata + NaN reductions for gw16")

# Same cell order in @meta.data and @reductions ?
neocortex_full@meta.data %>% rownames %>% tail
neocortex_full@reductions %>% map(~ tail(rownames(.)))

tic("Scale neocortex_full")
neocortex_full %<>% ScaleData(do.center = TRUE)  
push.toc()

tic("Convert")
seurat_to_h5ad(seurat_object = neocortex_full, 
               file_path_out = file.path("browser-dbx/data", 
                                     basename(neocortex_full_path) %>% 
                                       str_replace("rds", "h5ad")))
push.toc()
  

# Subset to try out scaling and converting to h5ad.
neocortex_full_subset <- subsetSeurat(object = neocortex_full, 
                                      fraction = 0.05, 
                                      write.to = "../neocortex/data/210322_neocortex_full_v3.2.3_subset_0.05.rds")

# Remove the extra columns that get added from subsetting
neocortex_full_subset@meta.data %<>% select(-nCount_RNA, -nFeature_RNA)
# Scale
neocortex_full_subset <- ScaleData(neocortex_full_subset)  
# Convert
seurat_to_h5ad(seurat_object = "../neocortex/data/210322_neocortex_full_v3.2.3_subset_0.05.rds", 
               file_path = "browser-dbx/data/210330_neocortex_v3.2.3_subset.05.h5seurat")
# WORKS. 1.5min for small file.
# there is no package called ‘SeuratObject’ Need to install SeuratObject for SeuratDisk functions.
# However SeuratObbject caused the neocortex_full merged object to become 4.0.0 which caused the error:
# Error in `[[<-`(`*tmp*`, dimreduc, value = tryCatch(expr = subset.DimReduc(x = x[[dimreduc]],  : 
#  [[<- defined for objects of type "S4" only for subclasses of environment.
```

## 211005

## Update to Schema 2.0:

[single-cell-curation/schema.md at main ·
chanzuckerberg/single-cell-curation](<https://github.com/chanzuckerberg/single-cell-curation/blob/main/schema/2.0.0/schema.md#development_stage_ontology_term_id>))

| **Key** | development_stage_ontology_term_id |
|---------|------------------------------------|

+-------------------+------------------------------------------------------------------------------+
| Fetal development | A term from the set of [9 to 38 week post-fertilization human                |
|                   | stages](http://www.ontobee.o%20%20%20%2                                      |
|                   | 0rg/search?ontology=HSAPDV&keywords=post-fertilization&submit=Search+terms)\ |
|                   | (9 weeks after conception and before birth; e.g.                             |
|                   | [HsapDv:0000046](http://purl.obolibrary.org/obo/HsapDv_0000046))             |
+-------------------+------------------------------------------------------------------------------+

```{r}
setwd("~/cortical_arealization/browser")

p_load(SeuratDisk)

# Load neocortex_full v3.2.3 object (contains gw16 and gw19)

neocortex <- read_rds("~/singlecell-neocortex-arealization/data/neocortex/210322_neocortex_full_v3.2.3.rds")

p_load(tictoc)
push.toc("Loaded neocortex")


```

## Add `developmental_stage_term_id` as a new column

Based on existing `age` column (gw).

```{r}
neocortex@meta.data %<>% mutate(age = str_glue_data(., "gw{age}"))

hsapdv_terms <- read_tsv(skip = 1, "cellxgene/browser-dbx/hsapdv_terms_to_gestational_weeks.tsv")

metadata_new <- neocortex@meta.data %>% dplyr::left_join(x = . , y = hsapdv_terms, 
                                          by = c("age" = "gestational_week")) %>%
  dplyr::select(age, developmental_stage_term_id, everything()) %>% 
  mutate(across(.cols = starts_with("percent"), .fns = round, 5))

neocortex@meta.data <- metadata_new %>% set_rownames(.$cell_id)
write_tsv(neocortex@meta.data, "cellxgene/browser-dbx/210322_neocortex_full_v3.2.3_metadata.tsv")

```

## Convert gene names to ENSEMBL_IDs

```{r}
p_load(biomaRt)

---
mart <- useMart("ensembl")
mart <- useDataset("hsapiens_gene_ensembl", mart = mart)
---
# Use GRCh37  
listEnsemblArchives()

biomaRt::listMarts()
biomaRt::listDatasets()

listMarts(host="https://grch37.ensembl.org")

mart <- useMart("ENSEMBL_MART_ENSEMBL", host = "https://grch37.ensembl.org")

searchDatasets(mart = mart, pattern = "hsapiens")

listDatasets(mart = mart)
```
```{r}
mart <- useDataset("hsapiens_gene_ensembl", mart = mart)
```

```{r}
attributes <- listAttributes(mart)

my_genes <- rownames(neocortex@assays$RNA@counts)

length(my_genes)
```

```{r}
 convertGenesToEnsembl <- function(my_values, my_filters) {
                                  
            lookup_table <- getBM(values = my_values,
                                  mart = mart,
                                  filters = my_filters,
                                  attributes = c("ensembl_gene_id",
                                  "external_gene_name",
                                  "external_gene_source",
                                  "hgnc_symbol",
                                  "external_synonym",
                                  "hgnc_id",
                                  "description",
                                  "gene_biotype"))
                        
            message(str_glue("\n • • • Matched {nrow(lookup_table)} out of {length(my_genes)}\n"))
                        
            eturn(lookup_table)
}
```

```{r}
externalname_to_ensembl <- convertGenesToEnsembl(my_values = my_genes, my_filters = "external_gene_name")

genesymbols_to_ensembl <- convertGenesToEnsembl(my_values = my_genes, my_filters = "hgnc_symbol")

synonym_to_ensembl <- convertGenesToEnsembl(my_values = my_genes, my_filters = "external_synonym")

```


externalname_to_ensembl %>% distinct(external_gene_name)


setdiff(my_genes, externalname_to_ensemblids$external_gene_name) %>% length
```{r}
my_genes %>% length
# 30,250

all_ensembl_matches <- bind_rows(list(external_name = externalname_to_ensembl,
                                      hgnc_symbol = genesymbols_to_ensembl,
                                      synonym = synonym_to_ensembl),
                                      .id = "match") %>%
                        distinct(ensembl_gene_id, .keep_all = TRUE )
# 30,841



externalname_to_ensembl %>% distinct(ensembl_gene_id)
# 30,637
genesymbols_to_ensembl %>% distinct(ensembl_gene_id)
# 21,651
synonym_to_ensembl %>% distinct(ensembl_gene_id)
# 2,082

all_ensembl_matches %<>%
    mutate(original_gene_name = case_when(match == "external_name" ~ external_gene_name,
                                           match == "hgnc_symbol" ~ hgnc_symbol,
                                           match == "synonym" ~ external_synonym))


setdiff(my_genes, all_ensembl_matches$original_gene_name) %>% length

# 2154
```

```{r}
genesymbols_to_ensembl <- list(original_gene_name = my_genes) %>%
                            as_tibble() %>%
                               left_join(all_ensembl_matches) %>%
  
    mutate(new_gene_id = case_when(is.na(ensembl_gene_id) ~ original_gene_name,
                                                     TRUE ~ ensembl_gene_id)) %>%
        dplyr::select(original_gene_name, new_gene_id, everything())

genesymbols_to_ensembl %<>% distinct(original_gene_name, .keep_all = TRUE)

write_tsv(genesymbols_to_ensembl, "cellxgene/browser-dbx/gene_symbols_to_ensembl_ids.tsv")

filter(genesymbols_to_ensembl, is.na(ensembl_gene_id))
# 2,154 missing ENSEMBL IDs

rownames(neocortex@assays$RNA@counts) <- genesymbols_to_ensembl$new_gene_id

write_rds(neocortex, "~/cortical_arealization/data/neocortex/2101006_neocortex_v3.2.3_ensembl_ids_schema2.0.rds")



RandomSubset<- function(object, rate, random.subset.seed = NULL, ...){
        ncells <- nrow(object@meta.data)
        ncells.subsample <- round(ncells * rate)

        set.seed(random.subset.seed)

        selected.cells <- sample(colnames(object), ncells.subsample)
        object <- subset(object, cells =  selected.cells,
                            ...)
        return(object)
}

small <- RandomSubset(neocortex, rate = 0.1 )

identical(colnames(neocortex@assays$RNA@counts), rownames(neocortex@meta.data))
TRUE
identical(colnames(neocortex@assays$RNA@data), rownames(neocortex@meta.data))
TRUE
identical(colnames(neocortex@assays$RNA@scale.data), rownames(neocortex@meta.data))
TRUE

SubsetData(neocortex, cells = )
write_rds(neocortex_subset, "neocortex/data/210318_neocortex_v3_subset.01.rds") 

neocortex@assays$RNA@counts %>% glimpse
```
```{r}
seurat2anndata(neocortex, outFile = "neocortex_full_v3.2.3_schema2.0_ensembl.hda5", main_layer = "scale.data", transfer_layers = "counts", drop_single_values = FALSE)
```


```{r}
neocortex_with_scale <- ScaleData(object = neocortex,
                                  do.center = TRUE, 
                                  block.size = 5000, verbose = TRUE, 
                                  features = rownames(data))  


save_rds(neocortex_with_scale, "~/cortical_arealization/data/neocortex/neocortex_full_v3.2.3.rds")
```
```{r}
library(SeuratObject)
library(SeuratData)

p_load(SeuratData)
data("pbmc3k.final")

small <- RandomSubset(pbmc3k.final, rate = 0.1)
small <- RandomSubset(pbmc3k, rate = 0.1)

pbmc3k %>% glimpse
```
```{r}
pbmc3k.final %>% glimpse
```
```{r}
pbmc3k.final@reductions %>% str(levels = 3)
```
```{r}
neocortex@reductions %>% str(levels = 3)
```
```{r}
neocortex@reductions %>% glimpse
```
```{r}
sessionInfo()
```

## Convert to h5ad


